---
title: "Sanbox: Explore FAVITES and custom contact matrix"
author: "Nicholas F. Brazeau"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    highlight: textmate
    theme: lumen
    toc: yes
    toc_float: yes
    toc_depth: 3
    code_folding: hide
editor_options: 
  chunk_output_type: console
---
```{r, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, eval = T, results = 'hide',
                      fig.align = 'center', fig.width = 6, fig.height = 6)
knitr::opts_knit$set(root.dir = here::here(""))

```

```{r}
library(tidyverse)
library(remotes)
remotes::install_github("nickbrazeau/goodegg"); library(goodegg)
remotes::install_github("CenterForAssessment/randomNames"); library(randomNames)
library(tidygraph)
library(ggraph)
library(cowplot)
library(ape)
library(treeio)
library(ggtree)
```
# Introduction/Purpose
1. Simulate a dataset that has individuals with differing contact patterns
2. Use a prediction approach to determine their "effective distance" 
- _i.e._ parameterize the distance matrix, undirected currently? 
3. Plug into `FAVITES` and run
4. **Q**: Can we recapture dyanmics of relationships in the genetic sequences 



# Simulate Relationships
Very simple framework: population of $N$ individuals that are a part of discrete demes, $K$ (e.g. spatial distance is discretized). There are $n_k$ individuals in each deme, such that $\sum_{k=1}^K n_k = N$. Individuals within a deme can be indexed with $j$, and each individual, $i_{j,k}$ has some predictor value that determines their frequency of contact that we will define as "magnetism", $p$. 


Assume that $p$ follows a beta distribution with higher values indicating more isolation and lower values indicating more sociality (_i.e._ this is an "amplifier" of distance... or more appropriately one is the loneliest number). Will assume Malecot's classic isolation by distance assumption and that "relatedness" (here contact) follows and exponential pattern.   
Demes are put into space on a square lattice model, where coordinates are inputted or predetermined, with euclidean distance dictating "distance" between them.
Then calculate an effective distance probability matrix and draw contacts from matrix to input into `FAVITES` (see notes for qualification of this). 


```{r}
#' @title Contact Probability Matrix Generator
#' @param k integer; number of demes to be considered
#' @param n integer vector; vector of length k with integer elements specifying 
#' number of individuals per deme
#' @param rbshape1 numeric; shape1 for the beta distribution to parameterize magnetism
#' @param rbshape2 numeric; shape2 for the beta distribution to parameterize magnetism
#' @param latdim integer; a value for the square lattice matrix for spatial 
#' coordinates to be randomly drawn from (mxm). Either it or coords must be specified 
#' @param coords matrix; x, y coordinates to be considered for each deme (assume) in same order as deme assignments
#' @details Magnetism beta distribution default is to have extreme tails -- some individuals are frequent isolaters vs frequent contactors. NB, parameterized that values closer to 1 are more isolator 
#' @details Remember, dealing with euclidean distance without standardization, so larger distances between nodes/demes will result in low contacts 
make_prob_contact_mat <- function(k, n, 
                                  rbshape1 = 0.5, rbshape2 = 0.5, 
                                  latdim = NULL, coords = NULL) {
  #......................
  # assertions
  #......................
  goodegg:::assert_int(k)
  goodegg:::assert_vector_int(n)
  goodegg:::assert_numeric(rbshape1)
  goodegg:::assert_numeric(rbshape2)
  if(!is.null(latdim)) {
    goodegg:::assert_int(latdim)
    # not assertion but to not waste if loop
    coords <- round(seq(1, latdim*latdim, by = latdim))
    coords <- expand.grid(coords, coords)
    coords <- coords[sample(1:nrow(coords), size = k), ]
  } else if (!is.null(coords)) {
    goodegg:::assert_ncol(coords, 2)
    goodegg:::assert_matrix_numeric(coords)
  } else {
    stop("Must specify either coords or latdim")
  }
  
  # calculate once
  N <- sum(n)
  
  #......................
  # calculate euclidean distance
  #......................
  distmat <- matrix(NA, k, k)
  diag(distmat) <- 1 # although distance is truly 0 w/in a deme, make 1 for identity of magnetism
  # get upper triangle (for loops are slower in R but for readability)
  ij <- t(combn(k,2))
  for(l in 1:nrow(ij)) {
    # get r index, remember r counts by rows first then columns
    e <- sum(ij[l,1] + (ij[l,2]-1)*k)
    distmat[e] <- dist(rbind(coords[ij[l,1], ], coords[ij[l,2], ]), 
                       method = "euclidean")
  }
  # euclidean distances are symmetric
  distmat[lower.tri(distmat)] <- t(distmat)[lower.tri(distmat)]
  
  #......................
  # draw sim/sample population
  #......................
  indp <- tibble::tibble(id = 1:N) %>% 
    dplyr::mutate(
      deme = rep(1:k, n),
      magn = rbeta(n = N, shape1 = rbshape1, shape2 = rbshape2)
    )
  
  #......................
  # ind contact
  #......................
  indmat <- matrix(NA, N, N)
  diag(indmat) <- 0 # assuming no one can truly have no contacts
  # loop through to fill in contact matrix (for loops are slower in R but for readability)
  ij <- t(combn(N,2))
  for(l in 1:nrow(ij)) {
    # get r index, remember r counts by rows first then columns
    e <- sum(ij[l,1] + (ij[l,2]-1)*N)
    # get params and dist
    m1 <- indp$magn[ij[l,1]]
    m2 <- indp$magn[ij[l,2]]
    d12 <- distmat[indp$deme[ij[l,1]], indp$deme[ij[l,2]]]
    indmat[e] <- exp(-( (m1+m2)*d12 ))
  }
  
  # standardize contact into a probability matrix 
  indmat[lower.tri(indmat)] <- t(indmat)[lower.tri(indmat)]
  return(indmat/rowSums(indmat))
}


#' @title Draw Contacts
#' @param cpm dataframe; contact prob matrix
#' @param iters interger; number of contacts to be generated 
#' @details assume a single seed of infection  
draw_contacts <- function(cpm, iters) {
  goodegg:::assert_matrix_numeric(cpm)
  goodegg:::assert_square_matrix(cpm)
  goodegg:::assert_single_int(iters)
  # storage
  contact_mat <- matrix(NA, nrow = iters, ncol = 2)
  #......................
  # draw contacts 
  #   cpm matrix is to->from setup 
  #......................
  infseed <- sample(nrow(cpm), size = 1)
  for (i in 1:iters) {
    # prop
    newcontact <- sample(1:ncol(cpm), size = 1, prob = cpm[infseed, ] )
    # update 
    contact_mat[i,] <- c(infseed, newcontact)
    infseed <- newcontact
  }
  
  # out
  colnames(contact_mat) <- c("id1", "id2")
  return(tibble::as_tibble(contact_mat))
}


#' @title Liftover to FAVITES format
#' @param outpath charpath; writing out to your local drive
#' @details adding random names purely for human readability later, is
#' an unneccessary intermed step

liftover_conmat2favites <- function(realized_contmat, outpath){
  rnnames <- randomNames::randomNames(n = length(unique(unlist(realized_contmat))),
                                      which.names = "first",
                                      sample.with.replacement = F)
  nodeval <- unique(unlist(realized_contmat))
  realized_contmat <- apply(realized_contmat, 2, as.character)
  for (i in 1:length(rnnames)) {
    realized_contmat[realized_contmat == nodeval[i]] <- rnnames[i]
  }
  
  # assert some class 
  #TODO 
  nodes <- sort(unique(as.vector(realized_contmat)))
  p1 <- data.frame(c1 = "NODE", c2 = nodes, c3 = ".", c4 = NA, c5 = NA)
  p2 <- data.frame(c1 = "EDGE", c2 = realized_contmat[,1], c3 = realized_contmat[,2],
                   c4 = ".", c5 = "u")
  # write out
  readr::write_tsv(x = rbind(p1,p2),
                   na = "",
                   col_names = F,
                   file = outpath)
  # return key for downstream 
  out <- tibble::tibble(node = 1:length(rnnames), names = rnnames)
  return(out)
}



```


## Run Functions
```{r}
contact_probmat <- make_prob_contact_mat(k = 5, n = rep(3,5),
                                     rbshape1 = 0.5, rbshape2 = 0.5,
                                     latdim = 5)

realized_contacts <- draw_contacts(cpm = contact_probmat, iters = 500) 
key <- liftover_conmat2favites(realized_contmat = realized_contacts,
                        outpath = "gears/play_contact_network.tab.txt")
```

## Run `FAVITES`
```{bash, engine.opts='-l', eval=T}
python3 engine/FAVITES/run_favites_docker.py --config gears/my_contact_config.json
sleep 5
```

## Viz Results
```{r}
#............................................................
# network
#...........................................................
# dictionary
dc1 <- key %>% 
  dplyr::rename(id1 = node) %>% 
  dplyr::mutate(id1 = as.character(id1))
dc2 <- key %>% 
  dplyr::rename(id2 = node) %>% 
  dplyr::mutate(id2 = as.character(id2))

# get edges in tidy format
adj_mat <- realized_contacts %>% 
  dplyr::mutate(uconnect = purrr::map2_chr(id1, id2,
                                           function(x,y){paste(sort(c(x,y)), collapse = "|")})) %>% 
  dplyr::group_by(uconnect) %>% 
  dplyr::summarise(
    conn_cnt = dplyr::n()
  ) %>% 
  dplyr::ungroup() %>% 
  dplyr::mutate(id1 = stringr::str_split_fixed(uconnect, "\\|", n=2)[,1],
                id2 = stringr::str_split_fixed(uconnect, "\\|", n=2)[,2]) %>% 
  dplyr::left_join(., dc1, by = "id1") %>% 
  dplyr::left_join(., dc2, by = "id2") %>% 
  dplyr::select(c("names.x", "names.y", "conn_cnt"))

# plot adjmatrix

p1 <- adj_mat %>% 
  tidygraph::as_tbl_graph(., directed = F) %>% 
  tidygraph::activate(., "nodes") %>%
  dplyr::mutate(community = as.factor(tidygraph::group_louvain(weights = conn_cnt))) %>%
  tidygraph::activate("edges") %>%
  ggraph::ggraph(layout = 'kk') +
  ggraph::geom_edge_link(aes(color = conn_cnt), width = 1) +
  ggraph::scale_edge_color_gradientn("Connections \n Count", colors = viridisLite::viridis(n = 15)) +
  ggraph::geom_node_point(aes(color = community), size = 3) +
  scale_color_brewer("Community", palette = "Set1") +
  geom_node_text(aes(label = name), repel = TRUE, size = 3)+
  ggraph::theme_graph() +
  theme(legend.position = "left") +
  guides(color = guide_legend(nrow = 2, byrow = TRUE))


#............................................................
# tree
#...........................................................
retseq <- ape::read.FASTA("results/mycontact_out_ret/error_free_files/sequence_data.fasta.gz")
names(retseq) <- stringr::str_extract(names(retseq), "\\|(.*?)\\|")
names(retseq) <- gsub(pattern = "\\|", replacement = "", x = names(retseq))

# going to assume a very simple model of evolution (and not test any)
Ndist <- ape::dist.dna(x = retseq, model = "N")
myTree <- nj(as.matrix(Ndist))

p2 <- ggtree(myTree, branch.length="none",  layout="circular") +
  geom_tiplab(size=3, color="#000000") +
  theme_tree()
```
```{r}
cowplot::plot_grid(p1,p2, nrow = 2, align = "h")
```


## Thoughts/Issues

- I've assumed this exponential decay of "contact". Even if we use fancy ML to
predict "magnetism", we sitll will have to parameterize that covariate. Would
need a full predictive matrix from the ML folks. Right now I have standardized 
the row probabilities to make a technically correct probability distribution for
each individual...but this is not in the spirit of $\beta$ in the WAIFW 
- Have assumed that there is no perfect isolation (_i.e._ no prob of contact)
- Have assumed undirected/symmetric contact probs (can't relax?)







